Illustrating splines using the Titanic dataset

### 2024-07-07 text/24/titanic-spline.Rmd
Updated 2024 Rmd files

---
title: Illustrating splines using the Titanic dataset
source: "New"
author: Steve Simon
date: 2024-05-07
categories:
- Blog post
tags:
- Nonlinear regression
- R programming
output: html_document
---

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(broom)
library(pspline)
library(readr)
library(rms)
library(tidyverse)
library(yaml)
f0 <- "https://raw.githubusercontent.com/pmean/datasets/master/"
f1 <- "titanic.yaml"
dd <- read_yaml(paste0(f0, f1))
```

@@

```{r}
f2 <- "titanic.txt"
ti0 <- 
  data.frame(
    read_tsv(
      file=paste0(f0, f2),
      col_types="ccncn"))
names(ti0) <- tolower(names(ti0))
head(ti0)
```

@@

```{r}
summary(ti0$Age)
count(ti0, pclass)
count(ti0, sex)
count(ti0, survived)
```

@@

```{r}
ggplot(data=ti0, aes(factor(survived), age)) + 
  geom_boxplot()
```

@@

```{r}
ti1 <- filter(ti0, !is.na(age))
linear_fit <- glm(
  survived ~ age,
  family=binomial,
  data=ti1)
tidy(linear_fit)
```

@@

```{r}
age_sequence <- data.frame(
  age = seq(min(ti1$age), max(ti1$age), length=100))
age_sequence$prediction <- 
  predict(linear_fit, newdata=age_sequence, type="response")
ggplot(data=ti1, aes(age, survived)) +
  geom_point(pch=1) +
  geom_line(
    data=age_sequence,
    aes(age, prediction))
```

@@

```{r}
age_sequence$prediction <- 
  predict(linear_fit, newdata=age_sequence, type="link")
ggplot(data=age_sequence, aes(age, prediction)) +
  geom_line()
```

@@

```{r}
spline_fit <- glm(
  survived ~ rcs(age),
  family=binomial,
  data=ti1)
tidy(spline_fit)
```

@@

```{r}
age_sequence$prediction <- 
  predict(spline_fit, newdata=age_sequence, type="link")
ggplot(data=age_sequence, aes(age, prediction)) +
  geom_line()
```

@@

```{r}
age_sequence$prediction <- 
  predict(spline_fit, newdata=age_sequence, type="response")
ggplot(data=ti1, aes(age, survived)) +
  geom_point(pch=1) +
  geom_line(
    data=age_sequence,
    aes(age, prediction))
```

>>>>

---
title: Illustrating splines using the Titanic dataset
source: "New"
author: Steve Simon
date: 2024-05-07
categories:
- Blog post
tags:
- Nonlinear regression
- R programming
- Rmarkdown graphs
output: html_document
page_update: complete
---

@@

```{r titanic-spline-01, echo=FALSE, message=FALSE, warning=FALSE}
library(broom)
library(pspline)
library(readr)
library(rms)
library(tidyverse)
library(yaml)
f0 <- "https://raw.githubusercontent.com/pmean/datasets/master/"
f1 <- "titanic.yaml"
dd <- read_yaml(paste0(f0, f1))
```

@@

```{r titanic-spline-02}
f2 <- "titanic.txt"
ti0 <- 
  data.frame(
    read_tsv(
      file=paste0(f0, f2),
      col_types="ccncn"))
names(ti0) <- tolower(names(ti0))
head(ti0)
```

@@

```{r titanic-spline-03}
summary(ti0$Age)
count(ti0, pclass)
count(ti0, sex)
count(ti0, survived)
```

@@

```{r titanic-spline-04}
ggplot(data=ti0, aes(factor(survived), age)) + 
  geom_boxplot()
```

@@

```{r titanic-spline-05}
ti1 <- filter(ti0, !is.na(age))
linear_fit <- glm(
  survived ~ age,
  family=binomial,
  data=ti1)
tidy(linear_fit)
```

@@

```{r titanic-spline-06}
age_sequence <- data.frame(
  age = seq(min(ti1$age), max(ti1$age), length=100))
age_sequence$prediction <- 
  predict(linear_fit, newdata=age_sequence, type="response")
ggplot(data=ti1, aes(age, survived)) +
  geom_point(pch=1) +
  geom_line(
    data=age_sequence,
    aes(age, prediction))
```

@@

```{r titanic-spline-07}
age_sequence$prediction <- 
  predict(linear_fit, newdata=age_sequence, type="link")
ggplot(data=age_sequence, aes(age, prediction)) +
  geom_line()
```

@@

```{r titanic-spline-08}
spline_fit <- glm(
  survived ~ rcs(age),
  family=binomial,
  data=ti1)
tidy(spline_fit)
```

@@

```{r titanic-spline-09}
age_sequence$prediction <- 
  predict(spline_fit, newdata=age_sequence, type="link")
ggplot(data=age_sequence, aes(age, prediction)) +
  geom_line()
```

@@

```{r titanic-spline-10}
age_sequence$prediction <- 
  predict(spline_fit, newdata=age_sequence, type="response")
ggplot(data=ti1, aes(age, survived)) +
  geom_point(pch=1) +
  geom_line(
    data=age_sequence,
    aes(age, prediction))
```
